---
title: "flchain"
output: html_document
date: "2023-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import the libraries
In those lines we import the library needed for the analysis: in particular, survival contains the dataset used; foreign is used for... ; survminer is used for...

```{r}
library(survival)
library(foreign)
library(survminer)


```


## Including Plots
In this section we present the plots of the analysis 
```{r}




```


### Task 1
In this section we performed the analysis for each covariate and make the estimation of the Cox regression model. 
 
```{r}

```

### Task 2
In this section we estimate the hazard ratio and the corresponding 95% confidence interval for the complete model
```{r}

```

### Task 3
Run the backward selection algorithm and check if it satisfies the PH assumption

```{r}

```




You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
